p-NLI: a Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-22 paper/application.tex
fix sentences refresh
add new files

\section{Put it All Together}
In this section, we utilizes the perturbation based analysis
to illustrate how the proposed system can be used
to diagnose the model behavior, study the model's attention mechanism.

@@

\subsection{What Does Attention Do?}

\subsection{Is Attention All You Need?}

>>>>

\section{Put it All Together}

\label{sec:caseStudy}

In this section, we utilizes the perturbation based analysis
to illustrate how the proposed system can be used
to diagnose the model behavior, study the model's attention mechanism.

@@

\subsection{How Attention Affect the Prediction?}

% \subsection{Is Attention All You Need?}

### 2018-03-22 paper/intro.tex
add new files

Recently, substantial research has been developed to combat the model interpretability challenge. Most notably, the works~\cite{SimonyanVedaldiZisserman2013, ZeilerFergus2014, YosinskiCluneNguyen2015, OlahMordvintsevSchubert2017, LiuShiLi2017, OlahSatyanarayanJohnson2018, BilalJourablooYe2018} on interpreting convolution neural network (CNN) model from both machine learning and visualization community have revealed the enormous potential and instant benefits of model interpretation, yet they also remind us how little we actually know about the inner mechanism of these deep neural networks.
%
Compare to image classification task, natural language processing systems have innate restrictions (e.g., the discrete nature of words) that makes the road to interpretation more treacherous. For example, \emph{feature visualization}~\cite{OlahMordvintsevSchubert2017}, an often deployed technique for illustrating what image features a given part of the network (e.g., neuron, layer, or channel) capture, can not be easily generalized to natural language. Image (pixel values) represents a continouous space, in which human can easily recognize the existance (or the absence) of patterns or shapes. For natural language, even though we can encode a word as a vector~\cite{MikolovSutskeverChen2013, PenningtonSocherManning2014}, the word embedding space is still discretely defined, where the interpolation between two vectors (i.e., two words) do not hold clear meaning.
%
These restrictions call for new avenues for solving the interpretation challenges of natural language models, which motivate the proposed work.

% Beside explore the internal states of the model, understand how different parts of the model work together and how one may affect one another is as or if not more important.
Moreover, despite many recent advances, the majority of the existing techniques study the model as an invariant object, where the model's behaviors are recorded and analyzed in an offline fashion.
%
However, the exploratory nature of the model interpretation analysis often leads to many \emph{what if ...} types of questions. Such as, what if I perturbed the current input? Will the prediction be stable? What if we change one of the critical internal states of the model, how would the modification affect the prediction? What if the current prediction is wrong, how and where should we change the model parameters to produce the correct result? These type of queries form a natural way to gain understanding and come up with hypotheses of the model mechanisms by \emph{interrogating} how the components of a model interact with each other in a dynamic setting. Theoretically, we can code up specialized experiments for each of these scenarios. However, such a process is not only tedious but also ignored the iterative nature of the exploratory analysis. Often, these questions are not pre-determined. Instead, new questions arise as we explore the data and analyze previous observations.

In this work, we aim to provide immediate and informative answers to these \emph{what if} questions by exploiting the expressive power of visualization and flexible integration with the underlying computation model. Instead of viewing the model as an invariant object, we approach the interpretability challenge by studying them in a dynamic environment. By employing a perturbation-driven paradigm, we probe the model and examine how changes in one part of the pipeline (the input, internal states, and output prediction, see Fig.~\ref{fig:modelPipeline}) affect others, which in turn provide a new perspective to attack the model interpretation problem.
%
We implement our vision of the perturbation-driven system around the natural language inference problem~\cite{Parikh2016}.
%
However, the componenets of the visualization and the overall paradigm can be easily extended to other NLP tasks, such as neural machine translation, and paragram summerization, etc.
%
In the simplest form, the inference task ask the question about whether \emph{sentence A} can infer the content of \emph{sentence B} (\emph{entail}), or they contradict with each other (\emph{contradiction}), or they talk about different/unrelated things (\emph{neutral}). Language inference addresses the fundimental challenge of sentence semantic similarity, and is one the core natural language process (NLP) tasks (detials in Section~\ref{sec:languageInference}).
%
One recent advance in NLP is the introduction of attention mechanisms~\cite{VaswaniShazeerParmar2017} (see details in Section~\ref{sec:attention}). Intuitively, the \emph{attention} captures where in the input the model should focus on, which assigns importance to individual word or a pair of word (i.e., the alignment between words in different sentences).
%
There has been many speculation on how attention works in various models. In this work, by utilizing the the proposed system, we explore how changes in sentence input, attention, or prediction affect each other. Such type of exploration is not only essentail to interpret the language inference problem, but also other NLP model utilized the same/similar mechansims.
%
Finally, we also enhanced the standard visual encoding (e.g., as a bipariate graph or as an matrix) of the attention matrix by overlaying sentence linguistic structure to allow grammar guided simplification and the ability to examine how attention corresponds to grammarical structure.
%

>>>>

Recently, substantial research has been developed to combat the model interpretability challenge. Most notably, the works~\cite{SimonyanVedaldiZisserman2013, ZeilerFergus2014, YosinskiCluneNguyen2015, OlahMordvintsevSchubert2017, LiuShiLi2017, OlahSatyanarayanJohnson2018, BilalJourablooYe2018} on interpreting convolution neural network (CNN) model from both machine learning and visualization community have revealed the enormous potential and instant benefits of model interpretation, yet they also remind us how little we actually know about the inner mechanism of these deep neural networks.

Compare to image classification task, natural language processing systems have innate restrictions (e.g., the discrete nature of words) that makes the road to interpretation more treacherous. For example, \emph{feature visualization}~\cite{OlahMordvintsevSchubert2017}, an often deployed technique for illustrating what image features a given part of the network (e.g., neuron, layer, or channel) capture, can not be easily generalized to natural language. Image (pixel values) represents a continuous space, in which human can easily recognize the existence (or the absence) of patterns or shapes. For natural language, even though we can encode a word as a vector~\cite{MikolovSutskeverChen2013, PenningtonSocherManning2014}, the word embedding space is still discretely defined, where the interpolation between two vectors (i.e., two words) do not hold clear meaning.
%
These restrictions call for new avenues for solving the interpretation challenges of natural language models, which motivate the proposed work.

% Beside explore the internal states of the model, understand how different parts of the model work together and how one may affect one another is as or if not more important.
Moreover, despite many recent advances, the majority of the existing techniques study the model as an invariant object, where the model's behaviors are recorded and analyzed in an offline fashion.
%
However, the exploratory nature of the model interpretation analysis often leads to many \emph{``what if ...''} types of questions. Such as, what if I perturbed the current input? Will the prediction be stable? What if we change one of the critical internal states of the model, how would the modification affect the prediction? What if the current prediction is wrong, how and where should we change the model parameters to produce the correct result? These type of queries form a natural way to gain understanding and come up with hypotheses of the model mechanisms by \emph{interrogating} how the components of a model interact with each other in a dynamic setting. Theoretically, we can code up specialized experiments for each of these scenarios. However, such a process is not only tedious but also ignored the iterative nature of the exploratory analysis. Often, these questions are not pre-determined. Instead, new questions arise as we explore the data and analyze previous observations.

In this work, we aim to provide immediate and informative answers to these \emph{``what if''} questions by combining the expressive power of visualization and direct online query of the neural network model. Instead of viewing the model as an invariant object, we approach the interpretability challenge by studying them in a dynamic environment. By employing a perturbation-driven paradigm, we probe the internal states of the model and examine how changes in one part of the pipeline (the input, internal states, and output prediction, see Fig.~\ref{fig:modelPipeline}) affect others, which in turn provide a new perspective to address the model interpretation problem.

We implement our vision of the perturbation-driven paradigm in an interactive visualization system for natural language inference model~\cite{Parikh2016}.
%
However, the components of the visualization and the overall paradigm can be easily extended to other NLP tasks, such as neural machine translation, and text summarization, etc.
%
In the simplest form, the inference task ask the question about whether \textbf{sentence A} can infer the content of \textbf{sentence B} (\emph{entail}), or they contradict with each other (\emph{contradiction}), or they talk about different/unrelated things (\emph{neutral}). Language inference addresses the fundamental challenge of measuring sentence semantic similarity, and is one the core natural language process (NLP) tasks (see detail in Section~\ref{sec:languageInference}).

One recent advance in neural natural language process models is the introduction of attention mechanisms~\cite{VaswaniShazeerParmar2017} (see Section~\ref{sec:attention}). Intuitively, the \emph{attention} captures where in the input the model determine it should focus on. The attention often represents as weights that assigned to individual word or a pair of word (i.e., the alignment between words in different sentences).
%
There has been many speculation on how attention works in various models. In this work, by utilizing the the proposed system, we are able to perturb the attention and observer the change in the prediction, perturb the sentence input and observe the effects on the attention. And more importantly, we also allow user to change prediction, where we apply an constrained optimization to find the least amount of parameter update to produce the user assign label. Another word, the user can explore how changes in sentence input, attention, or prediction affect each other. Such type of exploration is not only essential to interpret the language inference problem, but can also be easily generated to other NLP model utilized the attention mechanisms.
%
In addition, we also enhanced the standard visual encoding (e.g., as a bipartite graph or as an matrix) of the attention matrix by overlaying sentence linguistic structure to allow grammar guided simplification of the visual represent.
Finally, as discussed in Section~\ref{sec:caseStudy}, the ability to examine how attention corresponds to grammatical structure also help domain expert speculate the potential benefit of including sentence linguistic structure in the design of attention component of the model.
